High-order Solution Transfer between Curved Triangular Meshes

### 2024-12-01 doc/paper.tex
Remove commented out sections from `paper.tex`.
Made all edits from the reviewer.

%\newcommand{\rev}[1]{\textcolor{red}{#1}}
\newcommand{\rev}[1]{#1}

@@

\author{Per-Olof Persson}
\address{UC Berkeley, 970 Evans Hall \#3840, Berkeley, CA 94720-3840 USA}
\email{persson@berkeley.edu}
%% \urladdr{persson.berkeley.edu}

\begin{abstract}
\noindent The problem of solution transfer between meshes arises frequently in
computational physics, e.g. in Lagrangian methods where remeshing
occurs. The interpolation process must be conservative, i.e. it
must conserve physical properties, such as mass. We extend previous
works --- which described the solution transfer process for straight sided
unstructured meshes --- by considering high-order isoparametric meshes
with curved elements. To facilitate solution transfer, we numerically
integrate the product of shape functions via Green's theorem along the
boundary of the intersection of two curved elements. We perform a numerical
experiment and confirm the expected accuracy by transferring test fields
across two families of meshes.
\\ \\
\noindent \emph{Keywords}: Remapping, Curved Meshes, Lagrangian,
Solution Transfer, Discontinuous Galerkin
\end{abstract}

@@

Solution transfer between meshes is a common problem that
has practical applications to many methods in
computational physics. \rev{For example, by allowing the underyling
computational domain to change during a simulation, computational
effort can be focused dynamically to resolve relevant features
of a numerical solution. This so-called mesh adaptivity typically
requires translating the numerical solution from the old mesh to the new,
\rev{i.e. solution transfer \cite{Babuska1978,Peraire1987,Pain2001,Iske2004}}}. As another example, Lagrangian or particle-based
methods treat each node in the mesh as a particle and so with each timestep the
\rev{mesh travels \emph{with} the fluid \cite{MR3023731}.
However, over (typically short) time the mesh
becomes distorted and suffers a loss in element quality which eventually causes
inaccurate solutions.} To overcome this, the
domain must be remeshed or rezoned and the solution must be
transferred (remapped) onto the new mesh configuration. \rev{A related more general
class of methods is the Arbitrary Lagrangian-Eulerian (ALE) method, which
are not constrained by a fixed computational mesh (the Eulerian approach) or
by a fixed fluid flow (the Lagrangian approach) and typically combine the benefits
of both approaches \cite{Hirt1974,MR3302403}.}

Solution transfer is needed when a solution
(approximated by a discrete field) is known on a \emph{donor} mesh and must
be transferred to a \emph{target} mesh.
When pointwise interpolation is used to transfer a solution, quantities with
physical meaning (e.g. mass, concentration, energy) may not be conserved.
In many applications, the field
must be conserved for physical reasons, e.g. mass or energy cannot leave or
enter the system, hence this paper focuses on \emph{conservative} solution transfer
(typically using Galerkin or \(L_2\)-minimizing methods).

%% When pointwise interpolation is used to transfer a solution, quantities with
%% physical meaning (e.g. mass, concentration, energy) may not be conserved.
%% To address this, there have been many explorations of
%% \emph{conservative interpolation}. Since solution transfer is so commonly
%% needed in physical applications, this ...

The problem of conservative interpolation
has been considered previously for straight sided meshes.
However, both to allow for greater
geometric flexibility and for high order of convergence, this paper
addresses the case of curved isoparametric meshes. \rev{This is based on the
recent interest in so-called high-order methods \cite{Wang2013}, which have the ability to
produce highly accurate solutions with low dissipation and low dispersion error.
However, these methods typically require curved meshes to obtain the high accuracy.}

%% In addition, curved elements have advantages in problems that involve geometries that change over time,
%% such as flapping flight or fluid-structure interactions.

@@

%% Lagrangian methods treat each point in the
%% physical domain as a ``particle'' which moves along a characteristic curve
%% over time and then monitor values associated with the particle (heat / energy,
%% velocity, pressure, density, concentration, etc.). They are an effective way
%% to solve PDEs, even with higher order or non-linear terms \cite{MR3023731}.
%% This approach transforms the numerical solution of PDEs into a family of
%% numerical solutions to many independent ODEs. It allows the use of familiar
%% and well understood ODE solvers. In addition, Lagrangian methods often
%% have less restrictive conditions on time steps than Eulerian
%% methods\footnote{In Eulerian methods, the mesh is fixed.}.
%% When solving PDEs on unstructured meshes
%% with Lagrangian methods, the nodes move (since they
%% are treated like particles) and the mesh ``travels''.
%% A flow-based change to a mesh can cause problems if it
%% distorts the mesh until the element
%% quality is too low in some mesh elements. Over enough time, the mesh can
%% even tangle (i.e. elements begin to overlap).

%% To deal with distortion, one can allow the mesh to adapt in between time
%% steps. In addition to improving mesh quality, mesh
%% adaptivity can be used to dynamically focus computational effort to resolve
%% sensitive features of a numerical solution. From \cite{Iske2004}
%% \begin{quote}
%% {\small In order to balance the method's approximation quality and its
%% computational costs effectively, adaptivity is an essential requirement,
%% especially when modelling multiscale phenomena.}
%% \end{quote}
%% For more on mesh adaptivity, see \cite{Babuska1978, Peraire1987, Pain2001}.

%% In either case, the change in the
%% mesh between time steps requires transferring a known solution on the
%% discarded mesh to the mesh produced by the remeshing process. Without
%% the ability to change the mesh, Lagrangian methods (or, more generally,
%% ALE\footnote{ALE, or Arbitrary Lagrangian-Eulerian methods are not constrained
%% by a fixed computational mesh (the Eulerian approach) or a fixed fluid flow
%% (the Lagrangian approach) and typically combine the benefits of both approaches
%% to create composite methods} \cite{Hirt1974}) would not be useful, since after
%% a limited time the mesh will distort.

%% To allow for greater geometric flexibility and for high order of convergence,
%% curved mesh elements can be used in the finite element method. Though the
%% complexity of a method can steeply rise when allowing curved elements, the
%% trade for high-order convergence can be worth it. (See \cite{Wang2013} for
%% more on high-order CFD methods.) Curved meshes can typically
%% represent a given geometry with far fewer elements than a straight sided mesh.
%% The increase in accuracy also allows for the use of fewer elements, which
%% in turn can also facilitate a reduction in the overall computation time.

%% \begin{figure}
%%   \includegraphics{element_distortion.pdf}
%%   \centering
%%   \captionsetup{width=.75\linewidth}
%%   \caption{Movement of nodes in a quadratic element under distortion caused
%%     by particle motion along the velocity field \(\left[ y^2 \; 1 \right]^T\)
%%     from \(t = 0\) to \(t = 1\) with \(\Delta t = 1/2\). The green curves
%%     represent the characteristics that each node travels along.}
%%   \label{fig:element-distortion}
%% \end{figure}

%% Even if the domain has no inherent curvature, high-order (degree \(p\)) shape
%% functions allow for order \(p + 1\) convergence, which is desirable in
%% its own right. However, even in such cases, a Lagrangian method must either
%% curve the mesh or information about the flow of the geometry will be lost.
%% Figure~\ref{fig:element-distortion} shows what happens to a given
%% quadratic element as the nodes move along the
%% characteristics given by the velocity field
%% \(\left[ \begin{array}{c c} y^2 & 1 \end{array}\right]^T\).
%% This element uses the triangle vertices and edge
%% midpoints to determine the shape functions. However, as the nodes move
%% with the flow, the midpoints are no longer on the lines connecting
%% the vertex nodes. To allow the mesh to more accurately represent the
%% solution, the edges can instead curve so that the midpoint nodes remain
%% halfway between (i.e. half of the parameter space) the vertex nodes along
%% an edge.

%% In multiphysics simulations, a problem is partitioned into physical components.
%% This partitioning can apply to both the physical domain (e.g. separating a
%% solid and fluid at an interface) and the simulation solution itself (e.g.
%% solving for pressure on one mesh and velocity on another). Each (multi)physics
%% component is solved for on its own mesh. When the components interact, the
%% simulation solution must be transferred between those meshes.

%% In a similar category of application, solution transfer enables the comparison
%% of solutions defined on different meshes. For example, if a reference
%% solution is known on a very fine special-purpose mesh, the error can be
%% computed for a coarse mesh by transferring the solution from the fine
%% mesh and taking the difference. Or, if the same method is used on
%% different meshes of the same domain, the resulting computed solutions can
%% be compared via solution transfer. Or, if two different methods use two
%% different meshes of the same domain, a solution may need to be transferred
%% across these meshes for comparison.

Conservative solution transfer has been around since the advent of ALE,
and as a result much of the existing literature focuses on mesh-mesh
pairs that will occur during an ALE-based simulation. \rev{Many ALE methods
modify the mesh during simulation: when flow-based
mesh distortion occurs, elements are typically ``flipped'' (e.g. a
diagonal is switched in a pair of elements) or elements are subdivided
or combined.} These operations are inherently local, hence the solution
transfer can be done locally across known neighbors. Typically, this
locality is crucial to solution transfer methods. In \cite{Margolin2003},
the transfer is based on partitioning elements of the updated mesh into
components of elements from the old mesh and ``swept regions'' from
neighbouring elements. In \cite{Kucharik2008}, the (locally) changing
connectivity of the mesh is addressed. In \cite{Garimella2007}, the
local transfer is done on polyhedral meshes.

@@

\rev{In addition to conservation, there are many other considerations for a
solution transfer method that are important in various applications. For
example, since our focus is on high-order methods, these typically produce
oscillations in the solutions that can be non-physical and break the numerical
solvers. This can happen for the standard Galerkin projections that we
consider here, which means that e.g. discontinuities such as shocks in the
solution could become nonphysical after our solution transfer. Various
solutions to this have been proposed before, often referred to as limiting or
bound/monotonicity preserving methods, see e.g. \cite{MR3302403,Anderson2018}
where several approaches are studied including a high-order extension of the
flux-corrected transport (FCT) method. Other considerations for a solution
transfer scheme include the preservation of additional constraints, such
as the divergence or the curl of the solution.}

In this paper, an algorithm for conservative solution transfer between curved
meshes will be described. This method applies to meshes in \(\reals^2\).
Application to meshes in \(\reals^3\) is a direction for future research,
though the geometric kernels (see Chapter~\ref{sec:bezier-intersection})
become significantly more challenging to describe and implement. In
addition, the method will assume that every element in the
target mesh is contained in the donor mesh. This ensures that the solution
transfer is \rev{\emph{interpolative}}. In the case where all target elements are
partially covered, \emph{extrapolation} could be used to extend a solution
outside the domain, but for totally uncovered elements there is no clear
correspondence to elements in the donor mesh.

%%%%%%%%%%%%%%%
%% Overlapping meshes - exclude for now

%% \begin{figure}
%%   \includegraphics{main_figure29.pdf}
%%   \centering
%%   \captionsetup{width=.75\linewidth}
%%   \caption{Partially overlapping meshes on a near identical domain. Both are
%%     linear meshes that approximate the unit disc in \(\reals^2\). The outermost
%%     columns show how the domain of each mesh can be expanded so they agree.}
%%   \label{fig:partially-overlapping}
%% \end{figure}

%% The case of partially overlapping meshes can be addressed in particular
%% cases (i.e. with more information). For example, consider a problem
%% defined on \(\Omega = \reals^2\) and solution
%% that tends towards zero as points tend to infinity. A typical approach
%% may be to compute the solution on a circle of large enough radius and
%% consider the numerical solution to be zero outside the circle.
%% Figure~\ref{fig:partially-overlapping} shows how solution transfer could be
%% performed in such cases when the meshes partially overlap: construct a
%% simple region containing both computational domains and then mesh the
%% newly introduced area. However, the assumption that the numerical solution
%% is zero in the newly introduced area is very specific and
%% a similar approach may not apply in other cases of partial overlap.

%% Some attempts (\cite{Berger1987, Chesshire1994, Cai1999}) have been
%% made to interpolate fluxes between overlapping meshes. These perform
%% an interpolation on the region common to both meshes and then numerically
%% solve the PDE to determine the values on the uncovered elements.

%%%%%%%%%%%%%%%

%% When solving some PDEs with viscoelastic fields, additional nodes must be
%% tracked at quadrature points. As the mesh travels, these quadrature points
%% must be moved as well in a way that is consistent with the standard
%% nodes that define the element. The solution transfer method described here
%% does not provide a way either to determine quadrature points on a
%% target mesh or to map the field onto known quadrature points.

@@

We'll refer to \(\reals\) for the reals, \(\utri\) represents
the unit triangle (or unit simplex) in \(\reals^2\):
\begin{equation}
\utri = \left\{(s, t) \mid 0 \leq s, t, s + t \leq 1\right\}.
\end{equation}
When dealing with sequences with multiple indices, e.g.
\(s_{m, n} = m + n\), we'll use bold symbols to represent
a multi-index: \(\bm{i} = (m, n)\).
The binomial coefficient
\(\binom{n}{k}\) is equal to \(\frac{n!}{k! (n - k)!}\) and the trinomial
coefficient \(\binom{n}{i, j, k}\) is equal to \(\frac{n!}{i! j! k!}\)
(where \(i + j + k = n\)). The notation \(\delta_{ij}\) represents the
Kronecker delta, a value which is \(1\) when \(i = j\) and \(0\)
otherwise.

@@

A \emph{B\'{e}zier triangle} (\cite[Chapter~17]{Farin2001}) is a
mapping from the unit triangle
\(\utri\) and is determined by a control net
\(\left\{\bm{p}_{i, j, k}\right\}_{i + j + k = n} \subset \reals^d\).
A B\'{e}zier triangle is a particular kind of B\'{e}zier surface, i.e. one
in which there are two cartesian or three barycentric input parameters.
Often the term B\'{e}zier surface is used to refer to a tensor product or
rectangular patch.
For \((s, t) \in \utri\) we can define barycentric weights
\(\lambda_1 = 1 - s - t, \lambda_2 = s, \lambda_3 = t\) so that
\begin{equation}
1 = \left(\lambda_1 + \lambda_2 + \lambda_3\right)^n =
  \sum_{\substack{i + j + k = n \\ i, j, k \geq 0}} \binom{n}{i, j, k}
  \lambda_1^i \lambda_2^j \lambda_3^k.
\end{equation}
Using this we can similarly define a (triangular) Bernstein basis
\begin{equation}
B_{i, j, k}(s, t) = \binom{n}{i, j, k} (1 - s - t)^i s^j t^k
  = \binom{n}{i, j, k} \lambda_1^i \lambda_2^j \lambda_3^k
\end{equation}
that is in \(\left[0, 1\right]\) when \((s, t)\) is in \(\utri\).
Using this, we define points on the B\'{e}zier triangle as a
convex combination of the control net:
\begin{equation}
b(s, t) = \sum_{i + j + k = n} \binom{n}{i, j, k}
  \lambda_1^i \lambda_2^j \lambda_3^k
  \bm{p}_{i, j, k} \in \reals^d.
\end{equation}

@@

\noindent Rather than defining a B\'{e}zier triangle by the control net, it can
also be uniquely determined by the image of a standard lattice of
points in \(\utri\): \(b\left(j/n, k/n\right) = \bm{n}_{i, j, k}\);
we'll refer to these as \emph{standard nodes}.
Figure~\ref{fig:cubic-bezier-example} shows these standard nodes for
a cubic triangle in \(\reals^2\). To see the correspondence,
when \(p = 1\) the standard nodes \emph{are} the control net
\begin{equation}
b(s, t) = \lambda_1 \bm{n}_{1, 0, 0} +
\lambda_2 \bm{n}_{0, 1, 0} + \lambda_3 \bm{n}_{0, 0, 1}
\end{equation}
and when \(p = 2\)
\begin{multline}
b(s, t) = \lambda_1\left(2 \lambda_1 - 1\right) \bm{n}_{2, 0, 0} +
\lambda_2\left(2 \lambda_2 - 1\right) \bm{n}_{0, 2, 0} +
\lambda_3\left(2 \lambda_3 - 1\right) \bm{n}_{0, 0, 2} + \\
4 \lambda_1 \lambda_2 \bm{n}_{1, 1, 0} +
4 \lambda_2 \lambda_3 \bm{n}_{0, 1, 1} +
4 \lambda_3 \lambda_1 \bm{n}_{1, 0, 1}.
\end{multline}
However, it's worth noting that the transformation between
the control net and the standard nodes has condition
number that grows exponentially with \(n\) (see \cite{Farouki1991}, which
is related but does not directly show this).
This may make working with
higher degree triangles prohibitively unstable.

A \emph{valid} B\'{e}zier triangle is one which is
diffeomorphic to \(\utri\), i.e. \(b(s, t)\) is bijective and has
an \rev{invertible Jacobian everywhere}. We must also have the orientation
preserved, i.e. the Jacobian must have positive determinant. For example, in
Figure~\ref{fig:inverted-element}, the image of \(\utri\) under
the map \(b(s, t) = \left[\begin{array}{c c} (1 - s - t)^2 + s^2 & s^2 + t^2
\end{array}\right]^T\) is not valid because the Jacobian is zero along
the curve \(s^2 - st - t^2 - s + t = 0\) (the dashed line). Elements that
are not valid are called \emph{inverted} because they have regions with
``negative area''. For the example, the image \(b\left(\utri\right)\)
leaves the boundary determined by the edge curves: \(b(r, 0)\),
\(b(1 - r, r)\) and \(b(0, 1 - r)\) when \(r \in \left[0, 1\right]\).
This region outside the boundary is traced twice, once with
a positive Jacobian and once with a negative Jacobian.
\begin{figure}
  \includegraphics{inverted_element.pdf}
  \centering
  \captionsetup{width=.75\linewidth}
  \caption{The B\'{e}zier triangle given by \(b(s, t) = \left[
    (1 - s - t)^2 + s^2 \; \; s^2 + t^2 \right]^T\) produces an
    inverted element. It traces the same region twice, once with
    a positive Jacobian (the middle column) and once with a negative
    Jacobian (the right column).}
  \label{fig:inverted-element}
\end{figure}

@@

This fits a typical definition (\cite[Chapter~12]{FEM-ClaesJohnson})
of a curved element, but gives a special meaning to the mapping from
the reference triangle. Interpreting elements as B\'{e}zier triangles
has been used for Lagrangian methods where
mesh adaptivity is needed (e.g. \cite{CardozeMOP04}). Typically curved
elements only have one curved side (\cite{McLeod1972}) since they are used
to resolve geometric features of a boundary. See also
\cite{Zlmal1973, Zlmal1974}.
B\'{e}zier curves and triangles have a number of mathematical properties
(e.g. the convex hull property) that lead to elegant geometric
descriptions and algorithms.

Note that a B\'{e}zier triangle can be
determined from many different sources of data (for example the control net
or the standard nodes). The choice of this data may be changed to suit the
underlying physical problem without changing the actual mapping. Conversely,
the data can be fixed (e.g. as the control net) to avoid costly basis
conversion; once fixed, the equations of motion and other PDE terms can
be recast relative to the new basis (for an example, see \cite{Persson2009},
where the domain varies with time but the problem is reduced to
solving a transformed conservation law in a fixed reference configuration).

@@

When defining shape functions (i.e. a basis with geometric meaning) on a
curved element there are (at least) two choices. When the degree of the
shape functions is the same as the degree of the function being
represented on the B\'{e}zier triangle,
we say the element \(\mathcal{T}\) is \emph{isoparametric}.
For the multi-index
\(\bm{i} = (i, j , k)\), we define \(\bm{u}_{\bm{i}} =
\left(j/n, k/n\right)\) and the corresponding standard node
\(\bm{n}_{\bm{i}} = b\left(\bm{u}_{\bm{i}}\right)\).
Given these points, two choices for shape functions present
themselves:
\begin{itemize}
  \itemsep 0em
  \item \emph{Pre-Image Basis}:
    \(\phi_{\bm{j}}\left(\bm{n}_{\bm{i}}\right) =
      \widehat{\phi}_{\bm{j}}\left(\bm{u}_{\bm{i}}\right) =
      \widehat{\phi}_{\bm{j}}\left(b^{-1}\left(
      \bm{n}_{\bm{i}}\right)\right)\)
    where \(\widehat{\phi}_{\bm{j}}\) is a canonical basis function
    on \(\utri\), i.e.
    \(\widehat{\phi}_{\bm{j}}\) a degree \(p\) bivariate polynomial and
    \(\widehat{\phi}_{\bm{j}}\left(\bm{u}_{\bm{i}}\right) =
    \delta_{\bm{i} \bm{j}}\)
  \item \emph{Global Coordinates Basis}:
    \(\phi_{\bm{j}}\left(\bm{n}_{\bm{i}}\right) =
    \delta_{\bm{i} \bm{j}}\), i.e. a canonical basis function
    on the standard nodes \(\left\{\bm{n}_{\bm{i}}\right\}\).
\end{itemize}

>>>>

\newcommand{\rev}[1]{#1}
%%% TODO: remove \rev

@@

\author{Per-Olof Persson}
\address{UC Berkeley, 970 Evans Hall \#3840, Berkeley, CA 94720-3840 USA}
\email{persson@berkeley.edu}

\begin{abstract}
\noindent The problem of solution transfer between meshes arises frequently in
computational physics, e.g., in Lagrangian methods where remeshing
occurs. The interpolation process must be conservative, i.e., it
must conserve physical properties, such as mass. We extend previous
works --- which described the solution transfer process for straight sided
unstructured meshes --- by considering high-order isoparametric meshes
with curved elements. To facilitate solution transfer, we numerically
integrate the product of shape functions via Green's theorem along the
boundary of the intersection of two curved elements. We perform a numerical
experiment and confirm the expected accuracy by transferring test fields
across two families of meshes.
\\ \\
\noindent \emph{Keywords}: Remapping, Curved Meshes, Lagrangian,
Solution Transfer, Discontinuous Galerkin
\end{abstract}

@@

Solution transfer between meshes is a common problem that
has practical applications to many methods in
computational physics. \rev{For example, by allowing the underlying
computational domain to change during a simulation, computational
effort can be focused dynamically to resolve relevant features
of a numerical solution. This so-called mesh adaptivity typically
requires translating the numerical solution from the old mesh to the new,
\rev{that is, solution transfer \cite{Babuska1978,Peraire1987,Pain2001,Iske2004}}}.
As another example, Lagrangian or particle-based
methods treat each node in the mesh as a particle and so with each timestep the
\rev{mesh travels \emph{with} the fluid \cite{MR3023731}.
However, over (typically short) time the mesh
becomes distorted and suffers a loss in element quality which eventually causes
inaccurate solutions.} To overcome this, the
domain must be remeshed or rezoned and the solution must be
transferred (remapped) onto the new mesh configuration. \rev{A related more general
class of methods is the Arbitrary Lagrangian-Eulerian (ALE) method, which
are not constrained by a fixed computational mesh (the Eulerian approach) or
by a fixed fluid flow (the Lagrangian approach) and typically combine the benefits
of both approaches \cite{Hirt1974,MR3302403}.}

Solution transfer is needed when a solution
(approximated by a discrete field) is known on a \emph{donor} mesh and must
be transferred to a \emph{target} mesh.
When pointwise interpolation is used to transfer a solution, quantities with
physical meaning (for instance, mass, concentration, energy) may not be
conserved. In many applications, the field
must be conserved for physical reasons, e.g., mass or energy cannot leave or
enter the system; hence we focus on \emph{conservative} solution transfer
(typically using Galerkin or \(L_2\)-minimizing methods).

The problem of conservative interpolation
was considered previously for straight sided meshes.
However, both to allow for greater
geometric flexibility and for high order of convergence, we
address the case of curved isoparametric meshes. \rev{This is based on the
recent interest in so-called high-order methods \cite{Wang2013}, which have the ability to
produce highly accurate solutions with low dissipation and low dispersion error.
However, these methods typically require curved meshes to obtain the high accuracy.}

@@

Conservative solution transfer has been around since the advent of ALE,
and as a result much of the existing literature focuses on mesh-mesh
pairs that will occur during an ALE-based simulation. \rev{Many ALE methods
modify the mesh during simulation: when flow-based
mesh distortion occurs, elements are typically ``flipped'' (for example, a
diagonal is switched in a pair of elements) or elements are subdivided
or combined.} These operations are inherently local; hence the solution
transfer can be done locally across known neighbors. Typically, this
locality is crucial to solution transfer methods. In \cite{Margolin2003},
the transfer is based on partitioning elements of the updated mesh into
components of elements from the old mesh and ``swept regions'' from
neighboring elements. In \cite{Kucharik2008}, the (locally) changing
connectivity of the mesh is addressed. In \cite{Garimella2007}, the
local transfer is done on polyhedral meshes.

@@

\rev{In addition to conservation, there are many other considerations for a
solution transfer method that are important in various applications. For
example, since our focus is on high-order methods, these typically produce
oscillations in the solutions that can be non-physical and break the numerical
solvers. This can happen for the standard Galerkin projections that we
consider here, which means that e.g., discontinuities such as shocks in the
solution could become nonphysical after our solution transfer. Various
solutions to this have been proposed before, often referred to as limiting or
bound/monotonicity preserving methods; see, for instance,
\cite{MR3302403,Anderson2018}
where several approaches are studied including a high-order extension of the
flux-corrected transport (FCT) method. Other considerations for a solution
transfer scheme include the preservation of additional constraints, such
as the divergence or the curl of the solution.}

We describe an algorithm for conservative solution transfer between curved
meshes will be described. This method applies to meshes in \(\reals^2\).
Application to meshes in \(\reals^3\) is a direction for future research,
though the geometric kernels (see Chapter~\ref{sec:bezier-intersection})
become significantly more challenging to describe and implement. In
addition, the method will assume that every element in the
target mesh is contained in the donor mesh. This ensures that the solution
transfer is \rev{\emph{interpolative}}. In the case where all target elements are
partially covered, \emph{extrapolation} could be used to extend a solution
outside the domain, but for totally uncovered elements there is no clear
correspondence to elements in the donor mesh.

@@

We'll refer to \(\reals\) for the reals, \(\utri\) represents
the unit triangle (or unit simplex) in \(\reals^2\):
\begin{equation}
\utri = \left\{(s, t) \mid 0 \leq s, t, s + t \leq 1\right\}.
\end{equation}
When dealing with sequences with multiple indices, for example,
\(s_{m, n} = m + n\), we'll use bold symbols to represent
a multi-index: \(\bm{i} = (m, n)\).
The binomial coefficient
\(\binom{n}{k}\) is equal to \(\frac{n!}{k! (n - k)!}\) and the trinomial
coefficient \(\binom{n}{i, j, k}\) is equal to \(\frac{n!}{i! j! k!}\)
(where \(i + j + k = n\)). The notation \(\delta_{ij}\) represents the
Kronecker delta, a value which is \(1\) when \(i = j\) and \(0\)
otherwise.

@@

A \emph{B\'{e}zier triangle} \cite[Chapter~17]{Farin2001} is a
mapping from the unit triangle
\(\utri\) and is determined by a control net
\(\left\{\bm{p}_{i, j, k}\right\}_{i + j + k = n} \subset \reals^d\).
A B\'{e}zier triangle is a particular kind of B\'{e}zier surface, i.e., one
in which there are two cartesian or three barycentric input parameters.
Often the term B\'{e}zier surface is used to refer to a tensor product or
rectangular patch.
For \((s, t) \in \utri\) we can define barycentric weights
\(\lambda_1 = 1 - s - t, \lambda_2 = s, \lambda_3 = t\) so that
\begin{equation}
1 = \left(\lambda_1 + \lambda_2 + \lambda_3\right)^n =
  \sum_{\substack{i + j + k = n \\ i, j, k \geq 0}} \binom{n}{i, j, k}
  \lambda_1^i \lambda_2^j \lambda_3^k.
\end{equation}
Using this we can similarly define a (triangular) Bernstein basis
\begin{equation}
B_{i, j, k}(s, t) = \binom{n}{i, j, k} (1 - s - t)^i s^j t^k
  = \binom{n}{i, j, k} \lambda_1^i \lambda_2^j \lambda_3^k
\end{equation}
that is in \(\left[0, 1\right]\) when \((s, t)\) is in \(\utri\).
Using this, we define points on the B\'{e}zier triangle as a
convex combination of the control net:
\begin{equation}
b(s, t) = \sum_{i + j + k = n} \binom{n}{i, j, k}
  \lambda_1^i \lambda_2^j \lambda_3^k
  \bm{p}_{i, j, k} \in \reals^d.
\end{equation}

@@

\noindent Rather than defining a B\'{e}zier triangle by the control net, it can
also be uniquely determined by the image of a standard lattice of
points in \(\utri\): \(b\left(j/n, k/n\right) = \bm{n}_{i, j, k}\);
we'll refer to these as \emph{standard nodes}.
Figure~\ref{fig:cubic-bezier-example} shows these standard nodes form
a cubic triangle in \(\reals^2\). To see the correspondence,
when \(p = 1\) the standard nodes \emph{are} the control net
\begin{equation}
b(s, t) = \lambda_1 \bm{n}_{1, 0, 0} +
\lambda_2 \bm{n}_{0, 1, 0} + \lambda_3 \bm{n}_{0, 0, 1}
\end{equation}
and when \(p = 2\)
\begin{multline}
b(s, t) = \lambda_1\left(2 \lambda_1 - 1\right) \bm{n}_{2, 0, 0} +
\lambda_2\left(2 \lambda_2 - 1\right) \bm{n}_{0, 2, 0} +
\lambda_3\left(2 \lambda_3 - 1\right) \bm{n}_{0, 0, 2} + \\
4 \lambda_1 \lambda_2 \bm{n}_{1, 1, 0} +
4 \lambda_2 \lambda_3 \bm{n}_{0, 1, 1} +
4 \lambda_3 \lambda_1 \bm{n}_{1, 0, 1}.
\end{multline}
However, it's worth noting that the transformation between
the control net and the standard nodes has condition
number that grows exponentially with \(n\) (see \cite{Farouki1991}, which
is related but does not directly show this).
This may make working with
higher-degree triangles prohibitively unstable.

A \emph{valid} B\'{e}zier triangle is one which is
diffeomorphic to \(\utri\), that is, \(b(s, t)\) is bijective and has
an \rev{invertible Jacobian everywhere}. We must also have the orientation
preserved, that is, the Jacobian must have positive determinant. For example, in
Figure~\ref{fig:inverted-element}, the image of \(\utri\) under
the map \(b(s, t) = \left[\begin{array}{c c} (1 - s - t)^2 + s^2 & s^2 + t^2
\end{array}\right]^T\) is not valid because the Jacobian is zero along
the curve \(s^2 - st - t^2 - s + t = 0\) (the dashed line). Elements that
are not valid are called \emph{inverted} because they have regions with
``negative area''. For the example, the image \(b\left(\utri\right)\)
leaves the boundary determined by the edge curves: \(b(r, 0)\),
\(b(1 - r, r)\) and \(b(0, 1 - r)\) when \(r \in \left[0, 1\right]\).
This region outside the boundary is traced twice, once with
a positive Jacobian and once with a negative Jacobian.
\begin{figure}
  \includegraphics{inverted_element.pdf}
  \centering
  \captionsetup{width=.75\linewidth}
  \caption{The B\'{e}zier triangle given by \(b(s, t) = \left[
    (1 - s - t)^2 + s^2 \; \; s^2 + t^2 \right]^T\) produces an
    inverted element. It traces the same region twice, once with
    a positive Jacobian (the middle column) and once with a negative
    Jacobian (the right column).}
  \label{fig:inverted-element}
\end{figure}

@@

This fits a typical definition (\cite[Chapter~12]{FEM-ClaesJohnson})
of a curved element, but gives a special meaning to the mapping from
the reference triangle. Interpreting elements as B\'{e}zier triangles
has been used for Lagrangian methods where
mesh adaptivity is needed (e.g., \cite{CardozeMOP04}). Typically curved
elements only have one curved side (\cite{McLeod1972}) since they are used
to resolve geometric features of a boundary. See also
\cite{Zlmal1973, Zlmal1974}.
B\'{e}zier curves and triangles have a number of mathematical properties
(e.g., the convex hull property) that lead to elegant geometric
descriptions and algorithms.

Note that a B\'{e}zier triangle can be
determined from many different sources of data (for example, the control net
or the standard nodes). The choice of this data may be changed to suit the
underlying physical problem without changing the actual mapping. Conversely,
the data can be fixed (for instance, as the control net) to avoid costly basis
conversion; once fixed, the equations of motion and other PDE terms can
be recast relative to the new basis (for an example, see \cite{Persson2009},
where the domain varies with time but the problem is reduced to
solving a transformed conservation law in a fixed reference configuration).

@@

When defining shape functions (i.e., a basis with geometric meaning) on a
curved element there are (at least) two choices. When the degree of the
shape functions is the same as the degree of the function being
represented on the B\'{e}zier triangle,
we say the element \(\mathcal{T}\) is \emph{isoparametric}.
For the multi-index
\(\bm{i} = (i, j , k)\), we define \(\bm{u}_{\bm{i}} =
\left(j/n, k/n\right)\) and the corresponding standard node
\(\bm{n}_{\bm{i}} = b\left(\bm{u}_{\bm{i}}\right)\).
Given these points, two choices for shape functions present
themselves:
\begin{itemize}
  \itemsep 0em
  \item \emph{preimage basis}:
    \(\phi_{\bm{j}}\left(\bm{n}_{\bm{i}}\right) =
      \widehat{\phi}_{\bm{j}}\left(\bm{u}_{\bm{i}}\right) =
      \widehat{\phi}_{\bm{j}}\left(b^{-1}\left(
      \bm{n}_{\bm{i}}\right)\right)\)
    where \(\widehat{\phi}_{\bm{j}}\) is a canonical basis function
    on \(\utri\), that is,
    \(\widehat{\phi}_{\bm{j}}\) is a degree \(p\) bivariate polynomial and
    \(\widehat{\phi}_{\bm{j}}\left(\bm{u}_{\bm{i}}\right) =
    \delta_{\bm{i} \bm{j}}\).
  \item \emph{global coordinates basis}:
    \(\phi_{\bm{j}}\left(\bm{n}_{\bm{i}}\right) =
    \delta_{\bm{i} \bm{j}}\), that is, a canonical basis function
    on the standard nodes \(\left\{\bm{n}_{\bm{i}}\right\}\).
\end{itemize}

\noindent For example, consider a quadratic B\'{e}zier triangle:
\begin{gather}
b(s, t) = \left[ \begin{array}{c c}
    4 (s t + s + t) & 4 (s t + t + 1)
  \end{array}\right]^T \\
\Longrightarrow
\left[ \begin{array}{c c c c c c}
    \bm{n}_{2, 0, 0} &
    \bm{n}_{1, 1, 0} &
    \bm{n}_{0, 2, 0} &
    \bm{n}_{1, 0, 1} &
    \bm{n}_{0, 1, 1} &
    \bm{n}_{0, 0, 2}
  \end{array}\right] = \left[ \begin{array}{c c c c c c}
    0 & 2 & 4 & 2 & 5 & 4 \\
    4 & 4 & 4 & 6 & 7 & 8
  \end{array}\right].
\end{gather}
In the \emph{global coordinates basis}, we have
\begin{equation}
\phi^{G}_{0, 1, 1}(x, y) = \frac{(y - 4) (x - y + 4)}{6}.
\end{equation}
For the \emph{preimage basis}, we need the inverse
and the canonical basis
\begin{equation}
b^{-1}(x, y) = \left[ \begin{array}{c c}
    \frac{x - y + 4}{4} & \frac{y - 4}{x - y + 8}
  \end{array}\right] \quad \text{and} \quad
\widehat{\phi}_{0, 1, 1}(s, t) = 4 s t,
\end{equation}
and together they give
\begin{equation}
\phi^{P}_{0, 1, 1}(x, y) = \frac{(y - 4) (x - y + 4)}{x - y + 8}.
\end{equation}
In general \(\phi_{\bm{j}}^P\) may not even be a rational bivariate
function; due to composition with \(b^{-1}\) we can only guarantee that
it is algebraic (i.e., it can be defined as the zero set of polynomials
with coefficients in \(\reals\left[x, y\right]\)).
